Time and space specifications for stencils

### 2016-06-20 paper.tex
stripped out some of temporal parts

Figure~\ref{fig:syntax} gives the syntax of stencil specifications,
which we introduces in stages below.  The top-level is given by the
\nonterm{specification} production which splits into either a
\nonterm{regionDec} (region declaration) or a \nonterm{specDec}
(specification declaration). Specification declarations associate a
specification to one or more program variables $\bar{v}$ (here given
using the Fortran style syntax of $\ldots \; \texttt{::} \bar{v}$).
Specification declarations are either \nonterm{spatial} or
\nonterm{temporal} specifications and describe, for spatial
specifcations how an array variable $v$ is read, and for temporal
specifications what variables are used to define the array variable $v$.
%%

\subsubsection{Spatial specification syntax}

@@

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{D}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v
  \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\
\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}   \\[1em]
\nonterm{spatial} ::= & [\nonterm{mod}] \;
                       [\nonterm{approxMod},] \; [\nonterm{region}] \\
\nonterm{mod} ::= & [\nonterm{modR},] \; [\nonterm{modRefl},] \; [\nonterm{modIrefl},] \\
\nonterm{modR} ::= &
 \term{readOnce} \\
\nonterm{modRefl} ::= & \term{reflexive}, \; \term{dims=}\dimTy \{\term{,} \; \dimTy\} \\
\nonterm{modIrrefl} ::= & \term{irreflexive}, \; \term{dims=}\dimTy \{\term{,}
          \dimTy\}\\
\nonterm{approxMod} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \stenFwd{\mathbb{N}_{>0}}{\dimTy}  \\
\mid \; & \stenBwd{\mathbb{N}_{>0}}{\dimTy} \\
\mid \; & \stenCen{\mathbb{N}_{>0}}{\dimTy} \\
\mid \; & \nonterm{region} \; \term{+} \; \nonterm{region} \\
\mid \; & \nonterm{region} \; \term{*} \; \nonterm{region} \\
\mid \; & \nonterm{rvar}  \\[0.5em]
\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
  \\[0.5em]
\dimTy ::= \; & \mathbb{N}_{>0} \\
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\end{figure}

@@

In some cases, it is useful to give a lower and/or upper bound for a
stencil. This can be done using either the \term{atMost} or
\term{atLeast} modifiers. This is particularly useful in situations
where there is a non-contiguous stencil, which does not fit the rest
of the specification syntax.  For example:
%%
\begin{ExmVerbatim}
!= stencil atLeast, reflexive(dims=1)      :: a
!= stencil atMost, forward(depth=2, dim=1) :: a
b(i) = a(i) + a(i+2) 
\end{ExmVerbatim}
%%
\subsubsection{Temporal specifications}

@@

\begin{figure}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[1em]
\begin{array}{rll}
(\textsc{F\; \texttt{+} \;F}) \;\; &
\stenFwdS{n \, \textsf{max} \, m}{d} \\
 \equiv \; & \stenFwdS{n}{d} \; \texttt{+} \; \stenFwdS{m}{d} \\[1em]
%
(\textsc{B\; \texttt{+} \;B}) \;\; &
\stenBwdS{n \, \textsf{max} \, m}{d} \\
 \equiv \; & \stenBwdS{n}{d} \; \texttt{+} \; \stenBwdS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;C}) \;\; &
\stenCenS{n \, \textsf{max} \, m}{d} \\
\equiv \; & \stenCenS{n}{d} \; \texttt{+} \; \stenCenS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;F}) \;\; & \stenCenS{n}{d} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenS{n}{d} \; \texttt{+} \;
                      \stenFwdS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;B}) \;\; & 
\stenCenS{n}{d} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenS{n}{d} \; \texttt{+} \;
                      \stenBwdS{m}{d} \\[1em]
(\textsc{B\; \texttt{+} \;F}) \;\; & 
\stenCenS{n}{d} \\
\equiv \; & \stenFwdS{n}{d} \; \texttt{+} \; \stenBwdS{n}{d} \\[1em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em] 
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T) \\[0.5em] 
(\textsc{inter}) \;\; & (S \; \texttt{+} \; S') \; \texttt{*} \;
                              (T \; \texttt{+} \; T') \\
 \equiv \; & (S \;  \texttt{*} \; S) 
 \; \texttt{+} \; (S \; \texttt{*} \; T') 
 \; \texttt{+} \; (S' \; \texttt{*} \; T)
 \; \texttt{+} \; (S' \; \texttt{*} T')  \\[0.5em] 
\end{array} \\ \\
%
\framebox{${\textit{specDec} :: \texttt{v}} \equiv 
{\textit{specDec'} :: \texttt{v}}$} \\[1em]
\begin{array}{rl}
(\textsc{mutual}) \; & 
\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
  :: \texttt{w} 
\\
\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; & 
%\texttt{stencil} \; \texttt{dependency} (\bar{v}) 
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w}) 
%  :: \texttt{v} 
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
(\textsc{spatial}) \; &
(\textit{region} \equiv \textit{region'}) \\
\Rightarrow \; & (\texttt{stencil} \; [\textit{mod}] \; [\textit{approxMod},] \;
\textit{region} \\
& \!\!\equiv \; \texttt{stencil} \; [\textit{mod}] \;
            [\textit{approxMod},] \;
\textit{region}') \\
\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

\begin{figure}[t]
\begin{minipage}{0.88\linewidth}
\begin{align*}
\hspace{-0.7em}
\begin{array}{c}
\dfrac{}{S <: S}(\textsc{refl}) \qquad \dfrac{R <: S \quad S <: T}{R <:
  T}(\textsc{trans}) \\[1.5em]
\\[1.5em]
\dfrac{}{\texttt{read-once} \, S <: S}(\textsc{rep})
\\[1.5em]
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \hspace{3em}}
{\begin{array}{rl}
\stenFwdS{m}{d} & \stenFwdS{n}{d}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \stenBwdS{m}{d} & <: \stenBwdS{n}{d}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenS{m}{d} & <: \stenCenS{n}{d}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdS{m}{d} & <: \stenCenS{n}{d} \\
\wedge \; \stenFwdS{m}{d} & <: \stenCenS{n}{d} \\
\end{array}}
%{\footnotesize{(\textsc{cover})}}
%\hspace{-0.1em}
\end{array}
\end{align*}
\end{minipage}
\caption{Inequations on specifications}
\label{fig:inequations}
\end{figure}

>>>>

Figure~\ref{fig:syntax} gives the syntax of stencil specifications,
which we introduces in stages below.  The top-level is given by the
\nonterm{specification} production which splits into either a
\nonterm{regionDec} (region declaration) or a \nonterm{specDec}
(specification declaration). Specification declarations associate a
specification to one or more program variables $\bar{v}$ (here given
using the Fortran style syntax of $\ldots \; \texttt{::} \bar{v}$),
describing how the array variables $\bar{v}$ are acessed.
%Specification declarations are either \nonterm{spatial} or
%\nonterm{temporal} specifications and describe, for spatial
%specifcations,
%Specifications declarations describe 
%how an array variable $v$ is read
%, and for temporal
%specifications what variables are used to define the array variable $v$.
%%

%\subsubsection{Spatial specification syntax}

@@

\begin{figure}[t]
\begin{align*}
\def\arraystretch{1.2}
\setlength{\arraycolsep}{0.2em}
\newcommand{\dimTy}{\mathbb{D}}
\begin{array}{rl}
\nonterm{specification} ::= & \nonterm{regionDec} \mid \nonterm{specDec} \\
\nonterm{specDec} ::= & \term{stencil} \; \nonterm{spec} \;
                        \texttt{::} \; v
  \\
\nonterm{regionDec} ::= &  \texttt{region} \; \nonterm{rvar} \; \texttt{=} \;
                         \nonterm{region}\\[1em]
%\nonterm{spec} ::= & \nonterm{spatial} \mid \nonterm{temporal}   \\[1em]
\nonterm{spec} ::= & [\nonterm{mod}] \;
                       [\nonterm{approxMod},] \; [\nonterm{region}] \\
\nonterm{mod} ::= & [\nonterm{modR},] \; [\nonterm{modRefl},] \; [\nonterm{modIrefl},] \\
\nonterm{modR} ::= &
 \term{readOnce} \\
\nonterm{modRefl} ::= & \term{reflexive} \, (\term{dims=}\dimTy \{\term{,} \; \dimTy\}) \\
\nonterm{modIrrefl} ::= & \term{irreflexive} \, (\term{dims=}\dimTy \{\term{,}
          \dimTy\}) \\
\nonterm{approxMod} ::= & \term{atMost} \; \mid \; \term{atLeast} \\[0.1em]
\nonterm{region} ::= & \stenFwd{\mathbb{N}_{>0}}{\dimTy}  \\
\mid \; & \stenBwd{\mathbb{N}_{>0}}{\dimTy} \\
\mid \; & \stenCen{\mathbb{N}_{>0}}{\dimTy} \\
\mid \; & \nonterm{region} \; \term{+} \; \nonterm{region} \\
\mid \; & \nonterm{region} \; \term{*} \; \nonterm{region} \\
\mid \; & \nonterm{rvar}  \\[0.5em]
%\nonterm{temporal} ::= \; & \term{dependency} \; (v \; \{ , v \}) [, \texttt{mutual}]
%  \\[0.5em]
\dimTy ::= \; & \mathbb{N}_{>0} \\
\nonterm{rvar} ::= \; & [\text{\term{a}-\term{z}$\,$\term{A}-\term{Z}$\,$\term{0}-\term{9}}]+
\end{array}
\end{align*}
\caption{Specification syntax (EBNF grammar)}
\label{fig:syntax}
\end{figure}

@@

In some cases, it is useful to give a lower and/or upper bound for a
stencil. This can be done using either the \term{atMost} or
\term{atLeast} modifiers. This is particularly useful in situations
where there is a non-contiguous stencil, which does not fit the rest
of the specification syntax.  For example:
%%
\begin{ExmVerbatim}
!= stencil atLeast, reflexive(dims=1)      :: a
!= stencil atMost, forward(depth=2, dim=1) :: a
b(i) = a(i) + a(i+2) 
\end{ExmVerbatim}
%%
%\subsubsection{Temporal specifications}

@@

\begin{figure}
\begin{align*}
\setlength{\arraycolsep}{0.05em}
\begin{array}{c}
\framebox{$\textit{region} \equiv \textit{region}'$} \\[1em]
\begin{array}{rll}
(\textsc{F\; \texttt{+} \;F}) \;\; &
\stenFwdS{n \, \textsf{max} \, m}{d} \\
 \equiv \; & \stenFwdS{n}{d} \; \texttt{+} \; \stenFwdS{m}{d} \\[1em]
%
(\textsc{B\; \texttt{+} \;B}) \;\; &
\stenBwdS{n \, \textsf{max} \, m}{d} \\
 \equiv \; & \stenBwdS{n}{d} \; \texttt{+} \; \stenBwdS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;C}) \;\; &
\stenCenS{n \, \textsf{max} \, m}{d} \\
\equiv \; & \stenCenS{n}{d} \; \texttt{+} \; \stenCenS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;F}) \;\; & \stenCenS{n}{d} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenS{n}{d} \; \texttt{+} \;
                      \stenFwdS{m}{d} \\[1em]
%
(\textsc{C\; \texttt{+} \;B}) \;\; & 
\stenCenS{n}{d} \\
\textit{$m \leq n$} \; \equiv \; & \stenCenS{n}{d} \; \texttt{+} \;
                      \stenBwdS{m}{d} \\[1em]
(\textsc{B\; \texttt{+} \;F}) \;\; & 
\stenCenS{n}{d} \\
\equiv \; & \stenFwdS{n}{d} \; \texttt{+} \; \stenBwdS{n}{d} \\[1em]
%
(\texttt{+}\textsc{comm}) \;\; & S \; \texttt{+} \; T \; \equiv \; T \;
                       \texttt{+} \; S \\[0.5em]
%
(\texttt{+}\textsc{assoc}) \;\; & R \; \texttt{+} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{+} \; S) \; \texttt{+} \; T \\[0.5em]
(\texttt{*}\textsc{comm}) \;\; & S \; \texttt{*} \; T \; \equiv \; T \;
                       \texttt{*} \; S \\[0.5em]
%
(\texttt{*}\textsc{assoc}) \;\; & R \; \texttt{*} \; (S \; \texttt{*} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{*} \; T \\[0.5em] 
(\textsc{dist}) \;\; & R \; \texttt{*} \; (S \; \texttt{+} \; T) \; \equiv \; (R \;
                       \texttt{*} \; S) \; \texttt{+} \; (R
                       \; \texttt{*} \; T) \\[0.5em] 
(\textsc{inter}) \;\; & (S \; \texttt{+} \; S') \; \texttt{*} \;
                              (T \; \texttt{+} \; T') \\
 \equiv \; & (S \;  \texttt{*} \; S) 
 \; \texttt{+} \; (S \; \texttt{*} \; T') 
 \; \texttt{+} \; (S' \; \texttt{*} \; T)
 \; \texttt{+} \; (S' \; \texttt{*} T')  \\[0.5em] 
\end{array} \\ \\
%
\framebox{${\textit{specDec} :: \texttt{v}} \equiv 
{\textit{specDec'} :: \texttt{v}}$} \\[1em]
\begin{array}{rl}
% MUTUAL 
%(\textsc{mutual}) \; & 
%\texttt{stencil} \; \texttt{dependency} (\texttt{v}), \texttt{mutual}
%  :: \texttt{w} 
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\texttt{w}), \texttt{mutual} ::
%  \texttt{v} \\[0.5em]
%(\textsc{coalesce}) \; & 
%\texttt{stencil} \; \texttt{dependency} (\bar{v}) 
%  :: \texttt{v} \\
%& ; \texttt{stencil} \; \texttt{dependency} (\bar{w}) 
%  :: \texttt{v} 
%\\
%\equiv \; & \texttt{stencil} \; \texttt{dependency} (\bar{v}, \bar{w}) ::
%  \texttt{v} \\[0.5em]
(\textsc{spatial}) \; &
(\textit{region} \equiv \textit{region'}) \\
\Rightarrow \; & (\texttt{stencil} \; [\textit{mod}] \; [\textit{approxMod},] \;
\textit{region} \\
& \!\!\equiv \; \texttt{stencil} \; [\textit{mod}] \;
            [\textit{approxMod},] \;
\textit{region}') \\
\end{array}
\end{array}
\end{align*}
\caption{Equations on specifications}
\label{fig:equations}
\end{figure}

@@

\begin{figure}[t]
\begin{align*}
\begin{array}{c}
\dfrac{}{S <: S}(\textsc{refl}) \qquad \dfrac{R <: S \quad S <: T}{R <:
  T}(\textsc{trans}) \\[1.5em]
\dfrac{}{\texttt{read-once} \, S <: S}(\textsc{rep})
\\[1.5em]
\setlength{\arraycolsep}{0.1em}
\dfrac{\hspace{3em} m \leq n \hspace{3em}}
{\begin{array}{rl}
\stenFwdS{m}{d} & <: \stenFwdS{n}{d}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \stenBwdS{m}{d} & <: \stenBwdS{n}{d}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenCenS{m}{d} & <: \stenCenS{n}{d}
%& \;\, ds \subseteq es \wedge n \leq m
\\
\wedge \; \stenBwdS{m}{d} & <: \stenCenS{n}{d} \\
\wedge \; \stenFwdS{m}{d} & <: \stenCenS{n}{d} \\
\end{array}}
%{\footnotesize{(\textsc{cover})}}
%\hspace{-0.1em}
\end{array}
\end{align*}
\caption{Inequations on specifications}
\label{fig:inequations}
\end{figure}
